A Transactional Memory System for Parallel Python

### 2014-07-30 talk/dls2014/paper/paper.tex
be nicer to JS and update pdf

Dynamic languages like Python, PHP, Ruby, and JavaScript receive a lot
of attention but have not yet embraced parallelism. A parallel
programming model was not part of the design of those languages. Thus,
the reference implementations of, e.g., Python and Ruby use a single,
global interpreter lock (GIL) to serialise the execution of code in
threads. The use of a single global lock causes several disadvantages,
and as multi-core processors become the de facto standard in platforms
ranging from smart phones to data centres, the issue of parallel
execution must be addressed by these dynamic languages.

@@

This area can be seen as a continuous, segment-local array of bytes
that is indexed with an object's reference (SO) divided by 16. So
that each object has its own read marker, all objects have a size of
at least 16 bytes. Otherwise there could be false conflicts when
reading from two adjacent objects that share a single marker.
Instead of just setting the byte to \lstinline!true!  if the
corresponding object was read, we set it to a \lstinline!read_version!
belonging to the transaction, which will be incremented on each
commit.  Thereby, we can avoid resetting the bytes to
\lstinline!false!  on commit and only need to do this every 255
transactions. The whole code for the barrier shown in
Listing~\ref{lst:rb} is easily optimisable for compilers as well as
perfectly predictable for CPUs. Additionally, the read barrier is not
constrained to execute before the actual read -- both the compiler and
the CPU are free to reorder or group them for efficiency.

>>>>

Dynamic languages like Python, PHP, Ruby, and JavaScript receive a lot
of attention but only provide limited forms of parallelism. A parallel
programming model was not part of the initial design of those languages. Thus,
the reference implementations of, e.g., Python and Ruby use a single,
global interpreter lock (GIL) to serialise the execution of code in
threads. The use of a single global lock causes several disadvantages,
and as multi-core processors become the de facto standard in platforms
ranging from smart phones to data centres, the issue of parallel
execution must be addressed by these dynamic languages.

@@

This area can be seen as a continuous, segment-local array of bytes
that is indexed with an object's reference (SO) divided by 16. So
that each object has its own read marker, all objects have a size of
at least 16 bytes. Otherwise there could be false conflicts when
reading from two adjacent objects that share a single marker.
Instead of just setting the byte to \lstinline!true!  if the
corresponding object was read, we set it to a \lstinline!read_version!
belonging to the transaction, which will be incremented on each
commit.  Thereby, we can avoid resetting the bytes to
\lstinline!false!  on commit and only need to do this every 255
transactions. The whole code for the barrier shown in
Listing~\ref{lst:rb} is easily optimisable for compilers as well as
perfectly predictable for CPUs. Additionally, the read barrier is not
constrained to execute before the actual read -- both the compiler and
the CPU can reorder or group them freely between potential safe points
for additional performance.
